\section{Introdução}

\subsection{Context and Objectives}

\par This report details the work developed within the scope of the Multivariable, Non-linear, and Optimal Control course 
laboratory project. The primary objective is to design, simulate, and implement a robust control strategy to stabilize a highly 
non-linear and inherently unstable system using a pre-existing experimental setup of a rotary inverted pendulum.

\par The adopted methodology relies on linearizing the plant's mathematical model around its unstable equilibrium point 
(the upward vertical position) to derive a Linear Time-Invariant (LTI) state-space representation. Based on this model, a 
Linear Quadratic Gaussian (LQG) control architecture is developed. This comprises a Linear Quadratic Regulator (LQR) to compute 
the optimal state feedback law, and a Linear Quadratic Estimator (LQE, or Kalman Filter) to estimate the system's full internal 
state from noisy measurements. The project covers the theoretical analysis of the system, validation through MATLAB/Simulink 
simulations, and practical real-time implementation on the laboratory's physical hardware, specifically addressing real-world 
non-linearities such as motor deadzones and steady-state errors.

\subsection{System Description}

\par The physical plant utilized in this work is a rotary inverted pendulum apparatus (a Quanser experimental setup). 
The mechanical structure consists of a motorized rotary arm that actively maneuvers an unactuated, free-swinging pendulum rod 
attached to its extremity.

\vspace{1pt}

\par The system operates as a Single-Input Multiple-Output (SIMO) process. The single control input $u$ is the continuous 
voltage applied to the DC motor that drives the horizontal rotary arm. To monitor the system's behavior, two physical sensors 
provide the measured output vector $y$:

\vspace{1pt}

\begin{itemize}
    \item A potentiometer measuring the angular position of the rotary arm $\alpha$.
    \item An optical encoder measuring the precise angular position of the pendulum relative to the arm $\beta$.
\end{itemize}

\vspace{2pt}

\par Because the system's complete state vector includes both the angular positions $\alpha, \beta$ and their respective angular 
velocities $\dot{\alpha}, \dot{\beta}$, but only the positions are directly measured, the system requires a state observer. 
Furthermore, the provided physical setup interfaces with the Simulink control environment via A/D and D/A data acquisition boards, 
allowing the designed controllers to be executed and tested on the real system in real-time.